---
title: "PCA Analysis of Hip Hop Songs"
author: "eNVy"
date: '`r Sys.Date()`'
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)

library(tidyverse)
library(tidymodels)
library(spotifyr)
library(corrr)
```

Building a model using *tidyverse* and *tidymodels* with the use of Principal Component Analysis (PCA). Main goals are to understand the 

+ Song characteristics
+ Critic Scores associated with the songs

***************
The procedure will use; 

+ little bit of purrr
+ little bit of spotifyr
+ Major pre-processing
+ Major data handling
+ mild recipes

***************

## Explore the Data 

```{r}
# We'll be only using rankings data set.
rankings <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-14/rankings.csv") 

rankings %>% glimpse()
```

```{r}
rankings %>% ggplot(mapping = aes(year, points , color = gender)) + 
  geom_jitter( alpha = 0.5, ) + 
  scale_y_log10() + 
  theme_classic()
```
Notice the linear distribution of data between 1 ~ 20 of points. 

```{r}
# after obtainig the spotify access tokens from whatever developer account
# we can pull audio features from a created playlist
id <- '3bfec784c0c547ba8faf7e3af769277e'
secret <-  'b25faee713f84d03b56e09d0bc1bb5fd'
access_tokens <- get_spotify_access_token(client_id = id,client_secret = secret)
playlist <-  get_playlist_audio_features("tmock1923", "7esD007S7kzeSwVtcH9GFe", authorization = access_tokens)

playlist %>% head()
```
```{r}
playlist %>% count(time_signature) 
playlist %>% 
  filter(time_signature == 3) %>% 
  pull(track.name)
```

```{r}
pull_id <- function(query) {
  search_spotify(query, authorization = access_tokens, "track") %>% 
    arrange(-popularity) %>% 
    filter(row_number() == 1) %>%
    pull(id)
}


rankings %>% head(3) %>% 
  mutate(sQuery = paste(title, artist), 
         sQuery = str_to_lower(sQuery)) %>% 
  mutate(id =  map_chr(sQuery, pull_id))  %>% select(title, artist, id)

# the function fails when it pulls head(4)
# we can use possibly within purrr to identify where it goes wrong. 

rankings %>% head(5) %>% 
  mutate(sQuery = paste(title, artist), 
         sQuery = str_to_lower(sQuery)) %>% 
  mutate(id =  map_chr(sQuery, possibly(pull_id, NA_character_)))  %>% select(title, artist, id)

# it fails when searching the artist with ft. with some other artists
# so we'll remove everything after "ft." and then search

rankingID <- rankings %>%  
  mutate(sQuery = paste(title, artist), 
         sQuery = str_to_lower(sQuery), 
         sQuery = str_remove(sQuery, "ft.*$")) %>% 
  mutate(id =  map_chr(sQuery, possibly(pull_id, NA_character_)))
```

For `r scales::percent(mean(is.na(rankingID$id)))` percent of songs we can't find an spotify id. But that fine


```{r}
rankingFeatures <- rankingID %>% 
  mutate(idGroup  = row_number() %/% 80) %>% 
  select(idGroup, id) %>% 
  nest(data = c(id)) %>% 
  mutate(audioFeatures = map(data, ~get_track_audio_features(.$id, authorization = access_tokens)))
```


```{r}
rankingDF <- rankingID %>% 
  bind_cols(rankingFeatures %>% 
              select(audioFeatures) %>% 
              unnest(audioFeatures)) %>% 
  select(title, artist, points, year, danceability:tempo) %>% 
  na.omit()

rankingDF %>% head()
```


```{r}
rankingDF %>% 
  select(year:tempo) %>% 
  correlate() %>% 
  rearrange() %>% 
  shave() %>% 
  rplot() + silgelib::theme_plex()
```


```{r}
rankingLM <- rankingDF %>% 
  select(-title, -artist) %>% 
  lm(log(points) ~. ,data = .)

summary(rankingLM)
```

## Principal Component analysis

```{r}
hiphop_recipe <- recipe(points ~ . , data = rankingDF) %>% 
  update_role(title, artist, new_role = "id") %>% 
  step_log(points) %>% 
  step_normalize(all_predictors()) %>% 
  step_pca(all_predictors()) 

hiphop_rec_prep <- prep(hiphop_recipe)

```

### Explore the results of PCA

```{r}
tidy_pca <- yardstick::tidy(hiphop_rec_prep, 3)


tidy_pca %>% 
  mutate(component = fct_inorder(component)) %>% 
  ggplot(aes(x = value, y = terms, fill = terms)) + 
  geom_col(show.legend = FALSE) + 
  facet_wrap(~component) + 
  labs(y = NULL)
  
  
```

```{r}
tidy_pca %>% 
  filter(component %in% c("PC1", "PC2" , "PC3", "PC4")) %>% 
  group_by(component) %>% 
  top_n(6, abs(value)) %>% 
  ungroup() %>% 
  mutate(
    terms = tidytext::reorder_within(terms, abs(value), component)
  ) %>% 
  ggplot(aes(x = abs(value), y = terms, fill = value > 0)) + 
  geom_col() + 
  labs(x = "Absolute contribution to song popularity", y = NULL, fill = "Positive Contribution?") + 
  facet_wrap(~component, scales = "free_y") + 
  tidytext::scale_y_reordered()
  
```

```{r}
# fonts issue, might get stuck here. if so. load fonts loadfonts(win) or something
juice(hiphop_rec_prep) %>% 
  ggplot(aes(PC1, PC2, label = title))  + 
  geom_point(alpha = 0.2) + 
  geom_text(check_overlap = T, family = "Calibri Light")
```

```{r}
sdev = hiphop_rec_prep$steps[[3]]$res$sdev

percent_var = sdev ^ 2 / sum(sdev^2)

tibble(component = unique(tidy_pca$component), 
       percent_var = percent_var) %>% 
  mutate(
    component = fct_inorder(component)
  ) %>% 
  ggplot(aes(component, percent_var)) + 
  geom_col(show.legend = F, fill = "indianred4") + 
  scale_y_continuous(labels = scales::percent_format())

```


```{r}
pca_lm <- juice(hiphop_rec_prep) %>% 
  select(-title, -artist) %>% 
  lm(points ~., data = .)

summary(pca_lm)
```


## Conclusion 

**The model is LAAAMMMEEEE**

********

